---
title: "Reproducible Research Project 1"
author: "Alex Kaplan"
date: "October 15, 2015"
output: html_document
---

Rmarkdown file for Reproducible Research Project 1

Load and Process the Data

```{r, echo = TRUE}
setwd("~/R/DataScience/RepResearch")
library(RCurl)
library(knitr)
dataURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(dataURL, destfile ="activity.zip", method="curl")
unzip("activity.zip")
activity <- read.csv("activity.csv", header = TRUE)
head(activity)
str(activity)
#Convert the Date field to Date format
activity$date <- as.Date(as.character(activity$date), "%Y-%m-%d")
```

What is mean total number of activity taken per day?

```{r, echo = TRUE}
dailySteps <- aggregate(steps ~ date, data= activity, sum)
hist(dailySteps$steps, xlab = "Steps", ylab= "Frequency", main = "Step Histogram", breaks = 10)
stepMean <- mean(dailySteps$steps, na.rm=TRUE)
stepMean
stepMedian<- median(dailySteps$steps, na.rm=TRUE)
stepMedian
```

Mean is `r stepMean` and the Median is `r stepMedian`.

Average Daily Activity pattern
```{r, echo = TRUE}
intervalSteps <- aggregate(steps ~ interval, data = activity, mean)
plot(intervalSteps$interval,intervalSteps$steps,type="l", xlab= "Interval", ylab = "Average Steps", main= "Average Daily Steps by Interval")
intervalSteps[which.max(intervalSteps$steps),]
maxInterval <- intervalSteps[which.max(intervalSteps$steps),"interval"]
```
The 5 minute maximum interval is `r maxInterval`.

Imputing missing values
```{r, echo = TRUE}
blankRows <- sum(is.na(activity$steps))
blankRows
#substitute na's for mean from 5 minutes interval
activityWMeans <- activity
for (i in 1:nrow(activityWMeans)) {
  if (is.na(activityWMeans$steps[i])) {
    interval_value <- activityWMeans$interval[i]
    steps_value <- intervalSteps[
      intervalSteps$interval == interval_value,]
    activityWMeans$steps[i] <- steps_value$steps
  }
}
dailySteps2 <- aggregate(steps ~ date, data= activityWMeans, sum)
hist(dailySteps2$steps, xlab = "Steps", ylab= "Frequency", main = "Step Histogram Replaced NAs", breaks = 10)
stepMean2 <- mean(dailySteps2$steps, na.rm=TRUE)
stepMean2
stepMedian2<- median(dailySteps2$steps, na.rm=TRUE)
stepMedian2
```
There are `r blankRows` rows with NA.
The mean stayed the same, `r stepMean` and `r stepMean2`, but the median increased slightly `r stepMedian` and `r stepMedian2`.

Differences in activity patterns between weekdays and weekends.
```{r}
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
library(ggplot2)
?weekdays()
weekdays(activityWMeans$date[2000])
weekdays(activityWMeans$date[2000])%in% c("Sunday","Saturday")
activityWMeans$day <- weekdays(activityWMeans$date)
activityWMeans$type<- ifelse (activityWMeans$day %in% c('Sunday','Saturday'), 
       "Weekend", "Weekday")
activityWMeans$type <- as.factor(activityWMeans$type)
weekendSteps <- aggregate(steps ~ interval+type, data= activityWMeans, mean)
qplot(interval, steps, data=weekendSteps, type ='l', geom=c("line"), xlab= "Interval", ylab= "Number of Steps")+facet_wrap(~type, ncol=1)
```

